---
title: "STAT 308 -- Homework 5"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your \texttt{R} code with your answers, otherwise you will not be given full credit.  Please upload your assignment by Thursday, October 20, 11:59 pm in a pdf file to Sakai.

- 1. Consider the r dataset \texttt{X.rds}, which contains a $60\times4$ matrix of predictors and \texttt{y.rds} which contains a $60 \times 1$ vector of observed responses.

To load an .RDS file into \texttt{R}, use the function \texttt{readRDS} instead of \texttt{read.csv}

a. State the least squares regression line in the form $\hat{y} = \hat{\beta}_0 + \hat{\beta}_1x_1 + \hat{\beta}_2x_2 + \hat{\beta}_3x_3$
```{r}
```

b.  What is the sum of squared errors of the least squares regression line?
```{r}
```

c.  What is the estimated regression variance?
```{r}
```

d.  What is the estimate of the standard error for $\beta_1$?
```{r}
```

2.  Consider the dataset \texttt{economy.csv} which contains the following variables:

- \texttt{CRUDE}: dollars per barrel of crude oil
- \texttt{INTEREST}: interest on ten-year treasury notes
- \texttt{FOREIGN}: foreign investments in billions of dollars
- \texttt{DJIA}: Dow Jones industrial average
- \texttt{GNP}: Gross national product in billions of dollars
- \texttt{PURCHASE}: Purchasing power of U.S. dollar (in 1983 dollars)
- \texttt{CONSUMER}: Consumer debt in billions of dollars

Suppose we wish to create a linear model for crude oil price based on the other six variables in the dataset.

a.  State the least squares regression line for this linear model.
```{r}
```

b. Interpret the parameter associated with the variable \texttt{FOREIGN} in the context of the problem.
```{r}
```

c. Interpret the parameter associated with the variable \texttt{DJIA} in the context of the problem.
```{r}
```

d.  What are the error degrees of freedom for this linear model?
```{r}
```

e. What is the estimate of the regression variance?
```{r}
```

f. What are the sum of squared errors?
```{r}
```

g. State the value of $r^2$ and interpret this value in the context of the problem.
```{r}
```

