---
title: "STAT 308 -- Homework 4"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your \texttt{R} code with your answers, otherwise you will not be given full credit.  Please upload your assignment by Friday, February 17, 11:59 pm in an html/pdf/word file to Sakai.

- 1. Suppose we perform a simple linear regression with a sample size $n = 40$ where we obtain $$\bar{x} = -0.208, \bar{y} = 1.516, s_x = 2.354, s_y = 3.185$$ $$\hat{\beta}_0 = 1.745, \hat{\beta}_1 = 1.102,s_{\hat{\beta}_0} = 0.265, s_{\hat{\beta}_1} = 0.113$$

-- a. Calculate the sample correlation coefficient, $r$, and the $r^2$.
```{r}
beta1 <- 1.102
sx <- 2.354
sy <- 3.185
r <- beta1*sx/sy
r2 <- r^2
r
r2
```
$r = 0.8145$, $r^2 = 0.6634$

-- b. Calculate the estimate of the regression variance $s^2$. (Hint: Intuitively, this is the variance of $Y$ that is not explained through the linear model with $X$.)
```{r}
sy2 <- sy^2
n <- 40
SST <- sy2*(n-1)
SSE <- SST*(1-r2)
s2 <- SSE/(n-2)
s2
```
$s^2 = 3.505$

- 2. Use the following incomplete ANOVA table to answer the following questions.

|     |df    |Sums of Squares  |Mean Square  |F Value  |Pr(>f)   |
|-----|:-----|:----------------|:------------|:--------|:--------|
|Model|1 | 5.6             | | 3.3 |  |
|Error||  | |         |         |
|Total|40| |             |         |         |   

-- a. What is the mean squares for the model (MSM)?
```{r}
df_model <- 1
SSM <- 5.6
MSM <- SSM/df_model
MSM
```

$MSM = 5.6$

-- b. What is the mean squared error (MSE)?
```{r}
f <- 3.3
MSE <- MSM/f
MSE
```
$MSE = 1.697$

-- c. What are the error degrees of freedom?
```{r}
df_total <- 40
df_error <- df_total - df_model
df_error
```
$df_error = 39$

-- d. What is the sum of squared errors (SSE)?
```{r}
SSE <- MSE*df_error
SSE
```
$SSE = 66.18$

-- e. What is the p-value used to test for a significant linear relationship between $X$ and $Y$?
```{r}
p <- pf(f,df1=df_model,df2=df_error,lower.tail=FALSE)
p
```
$p = 0.0770$

- 3. Reconsider the dataset \texttt{AdRevenue.csv} as well as our simple linear regression model of ad revenue (in millions of dollars) based on circulation (in millions).

-- a. Obtain an ANOVA table for this model.  
```{r}
adrev <- read.csv("../Data/adrevenue.csv")
mod <- lm(AdRevenue ~ Circulation,adrev)
anova(mod)
```

-- b. Using the ANOVA table, perform an F test for a linear relationship between ad revenue and circulation.  Be sure to properly state your hypotheses, your test statistic, p-value, and a decision and conclusion at $\alpha = 0.05$.

$$H_0: \beta_1 = 0$$

$$H_a: \beta_1 \neq 0$$

```{r}
f <- anova(mod)[1,"F value"]
p <- anova(mod)[1,"Pr(>F)"]
f
p
```

$F = 576.5$, $p = 6.293 \times 10^{-35}$,  $p < \alpha$, therefore we reject $H_0$ and conclude that there is a statistically significant linear relationship between ad revenue and circulation.

-- c. What is the distribution the test statistic follows under $H_0$?  In other words, what is the distribution we use to calculate the p-value?

$F$ is an F-distribution with 1 and 68 degrees of freedom

-- d. Using the ANOVA table, calculate the value of $r^2$.  Interpret this value in the context of the problem.
```{r}
SSM <- anova(mod)[1,"Sum Sq"]
SSE <- anova(mod)[2,"Sum Sq"]
r2 <- SSM/(SSM + SSE)
r2
p <- anova(mod)[1,"Pr(>F)"]
```
89.45% of the variation in magazine's ad revenue can be explained through its linear relationship with number of magazines in circulation.

