```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your \texttt{R} code with your answers, otherwise you will not be given full credit.  Please upload your assignment by Friday, March 17, 11:59 pm in an html,pdf,or word file to Sakai.

1. Use the below degrees of freedom and sums of squares from ANOVA tables from a multiple linear regression with three explanatory variables to answer the following questions.

- Model 1: $X_1$ as only explanatory variable

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |1 | 2560.45 |
|Error | 58 | 251.08 |

- Model 2: $X_1$ and $X_2$ as explanatory variables

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |2 | 2733.63 |
|Error | 57 |77.90 |

- Model 3: $X_1$, $X_2$, and $X_3$ as explanatory variables

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |3 | 2780.43 |
|Error | 56 |31.1 |

a. What are the additional model sums of squares when we add $X_2$ to a model that includes $X_1$?

b. What is the reduction in the sum of squared errors when we add $X_2$ and $X_3$ to a model that includes $X_1$?

c. What is the mean squared error for the linear model with $X_1$, $X_2$, and $X_3$ as explanatory variables?

d. Find an F statistic and p-value for testing for overall significance of the linear model with $X_1$, $X_2$, and $X_3$ as explanatory variables.

e. Find an F stastistic and p-value for testing if adding $X_2$ and $X_3$ to a linear model that already includes $X_1$ is statistically significant.

2. Consider the dataset \texttt{economy.csv} which contains the following variables:

- \texttt{CRUDE}: dollars per barrel of crude oil
- \texttt{INTEREST}: interest on ten-year treasury notes
- \texttt{FOREIGN}: foreign investments in billions of dollars
- \texttt{DJIA}: Dow Jones industrial average
- \texttt{GNP}: Gross national product in billions of dollars
- \texttt{PURCHASE}: Purchasing power of U.S. dollar (in 1983 dollars)
- \texttt{CONSUMER}: Consumer debt in billions of dollars

a. Suppose we create a linear model with crude oil in dollars per barrel as the response variable and all of the other variables in the dataset as the explanatory variables.  Perform an overall F test to determine if this overall linear model is statistically significant.  Be sure to include all necessary information to perform a hypothesis test.  Assume $\alpha = 0.05$.

```{r}
```

b. Check if the assumptions of homoscedasticity and normally distributed residuals are violated for the model in (a).

```{r}
```

c.  Suppose we know that dollars per barrel is related linearly with interest rates.

(i).  Report and interpret the $r^2$ value for the linear model for dollars per barrel with interest rate as the only explanatory variable.
```{r}
```

(ii). Perform a hypothesis test to determine if adding purchasing power of the U.S. dollar to our linear model significantly improves the predictive ability of our model.  Be sure to include all necessary information to perform a hypothesis test.  Assume $\alpha = 0.05$.
```{r}
```

(iii). What is the additional percent variation in crude oil dollars per barrel than can be explained by adding purchasing power of U.S. dollar to our model to our linear model that already includes interest rates (Hint: this is the difference in the $r^2$s of the two models). 
```{r}
```

(iv). Perform a hypothesis test to determine if adding both purchasing power of the U.S. dollar and Dow Jones Industrial Average to our linear model that already includes interest rates significantly improves the predictive ability of our model.  Be sure to include all necessary information to perform a hypothesis test.  Assume $\alpha = 0.05$.
```{r}
```

