---
title: "STAT 308 -- Homework 2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your \texttt{R} code with your answers, otherwise you will not be given full credit.  Please upload your assignment by Thursday, September 15, 11:59 pm in a pdf file to Sakai.

- 1. In a simple linear regression problem where $n=30$, we obtain $$\sum_{i=1}^n x_i = 75, \sum_{i=1}^n y_i = 660, \sum_{i=1}^n x_i^2 = 240, \sum_{i=1}^n y_i^2 = 18000, \sum_{i=1}^n x_iy_i = 1200.$$

-- a. Calculate $\hat{\beta}_0$ and $\hat{\beta}_1$.

```{r}
# Write your code below this line
```

-- b. Calculate SSE and $s^2$.

```{r}
# Write your code below this line
```

- 2. Consider the dataset \texttt{AdRevenue.csv} on the course webpage.  Suppose we are interested in using the number of magazines in circulation (in millions) to predict magazine's ad revenue (in millions of dollars).

-- a. Draw a scatterplot of AdRevenue vs. Circulation.  Do you think a linear relationship between Circulation and AdRevenue is appropriate?  Justify your response.
```{r}
# Write your code below this line
```

-- b. Using \texttt{R}, find the equation of the least squares regression line. 
```{r}
# Write your code below this line
```

-- c. Add the least squares regression line to the scatterplot in (a).
```{r}
# Write your code below this line
```

-- d. Interpret the slope of the regression line in the context of the given problem.

-- e. Interpret the intercept of the regression line in the context of the given problem.  Does this interpretation make sense?  Why or why not?

-- f. What do we expect the amount of ad revenue to be when there are 4 million magazines in circulation?
```{r}
# Write your code below this line
```

-- g. Find the value of SSE and $s^2$ for the least squares regression line.
```{r}
# Write your code below this line
```

-- h. Determine if the assumption of homoscedasticity is violated.
```{r}
# Write your code below this line
```

-- i. Determine if the assumption of normally distributed residuals is violated.
```{r}
# Write your code below this line
```

