---
title: "STAT 308 -- Homework 4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your \texttt{R} code with your answers, otherwise you will not be given full credit.  Please upload your assignment by Thursday, September 29, 11:59 pm in a pdf file to Sakai.

- 1. Suppose we perform a simple linear regression where $$n=50, \bar{x} = -0.208, \bar{y} = 1.516, s_x = 2.354, s_y = 3.185$$ $$\hat{\beta}_0 = 1.745, \hat{\beta}_1 = 1.102, s_{Y|X} = 1.868,s_{\hat{\beta}_0} = 0.265, s_{\hat{\beta}_1} = 0.113$$

-- a. Calculate the sample correlation coefficient, $r$, and the $r^2$.
```{r}
```

-- b. Calculate the estimate of the regression variance $s_{y|x}^2$. (Hint: Intuitively, this is the variance of $Y$ that is not explained through the linear model with $X$.)
```{r}
```

- 2. Use the following incomplete ANOVA table to answer the following questions.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrr}
  \hline
 & df & Sum Sq & Mean Sq & F value & Pr($>$F) \\ 
  \hline
Model & 1 & 1.47 & & 0.18 &  \\ 
  Error & &  &  &  &  \\ 
   \hline
   Total & 34 & & & & \\
\end{tabular}
\end{table}

-- a. What is the mean squares for the model (MSM)?
```{r}
```

-- b. What is the mean squared error (MSE)?
```{r}
```

-- c. What are the error degrees of freedom?
```{r}
```

-- d. What is the sum of squared errors (SSE)?
```{r}
```

-- e. What is the p-value used to test for a significant linear relationship between $X$ and $Y$?
```{r}
```

- 3. Reconsider the dataset \texttt{AdRevenue.csv} as well as our simple linear regression model of ad revenue (in millions of dollars) based on circulation (in millions).

-- a. Obtain an ANOVA table for this model.  
```{r}
```

-- b. Using the ANOVA table, perform a hypothesis test for a linear relationship between ad revenue and circulation.  Be sure to properly state your hypotheses, your test statistic, p-value, and a decision and conclusion at $\alpha = 0.05$.
```{r}
```

-- c. What is the distribution the test statistic follows under $H_0$?  In other words, what is the distribution we use to calculate the p-value?

-- d. Using the ANOVA table, calculate the value of $r^2$.  Interpret this value in the context of the problem.
```{r}
```


