---
title: "STAT 308 -- Homework 1"
author: Solutions
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

- 1.
Assume that $Z$ is a random variable from the standard normal (i.e. $\mu = 0$, $\sigma = 1$).

-- a.
Calculate $P(Z \geq -1)$.
```{r}
# Write your R code to answer the question here.
pnorm(-1,mean=0,sd=1,lower.tail=FALSE)
```

OR 

```{r}
# Write your R code to answer the question here.
1 - pnorm(-1,mean=0,sd=1)
```

-- b.
Determine the $0.2-$quantile of $Z$.
```{r}
# Write your R code to answer the question here.
qnorm(0.2,mean=0,sd=1)
```

- 2.
$F_{a,b}$ is a random variable from the $F$-distribution with $df1 = a$ and $df2 = b$.

-- a.
Calculate $P(F_{5,40} \leq 2.9)$
```{r}
# Write your R code to answer the question here.
pf(2.9,df1=5,df2=40)
```
-- b.
Find $x$ such that $P(F_{6,24} \geq x) = 0.05$
```{r}
# Write your R code to answer the question here.
qf(0.05,df1=6,df2=24,lower.tail=FALSE)
```

OR

```{r}
# Write your R code to answer the question here.
qf(1-0.05,df1=6,df2=24)
```

- 3.
Given a dataset of scores $\{0,2,5,6,3,3,3,1,4,3\}$, calculate the

-- a.
mean,
```{r}
# Write your R code to answer the question here.
x <- c(0,2,5,6,3,3,3,1,4,3)
mean(x)
```
-- b.
median,
```{r}
# Write your R code to answer the question here.
median(x)
```
-- c.
variance.
```{r}
# Write your R code to answer the question here.
var(x)
```

- 4.
A random sample of 32 persons attending a certain diet clinic was found to have lost an average of 30 pounds over a three week period, with a sample standard deviation of 11 pounds. For these data

-- a.
Calculate a 99% confidence interval for the given data.
```{r}
# Write your R code to answer the question here.
n <- 32
est <- 30
se <- 11/sqrt(n)
alpha <- 0.01
crit <- qt(1-alpha/2,df=n-1)
est + c(-1,1)*crit*se
```
-- b.
Interpret the confidence interval in the context of the given problem.

We are 99% confident that the true mean weight loss over the last three weeks of all patients at the clinic is between 24.66 and 35.34 pounds.

-- c.
Suppose I wished to test my current belief that the average weight loss of the population is equal to 28 pounds.  I come to the conclusion to reject $H_0: \mu = 28$.  Based on your answer to (b), does this make sense?  Why or why not?

This does not make sense because 28 is inside our 99% confidence interval, which means that we would fail to reject $H_0$.

- 5.
An outbreak of Salmonella-related illness was thought to be caused by pre-cut melons from a specific factory.  Several samples were collected and can be found in the file \texttt{salmonella.csv} on the course webpage.  A Salmonella level (in MPN/g) greater than 0.3 MPN/g is considered dangerous. To demonstrate that the levels are safe we want to show the mean level is less than 0.3 MPN/g.

-- a.
State the null and directional alternative hypothesis in symbols.  Explicitly define $\mu$.

$H_0: \mu \leq 0.3$, $H_a: \mu > 0.3$.  $\mu$ is the average salmonella level (in MPN/g) of all melons from this factory.

-- b.
Find the test statistic and p-value for the data below.  Be sure to clearly identify them from your output.  Note you are performing a one-sided test.
```{r}
# Write your R code to answer the question here.
salmonella <- read.csv("../Data/salmonella.csv")
t.test(salmonella$Level,mu=0.3,alternative = "greater")
```

$t = -0.1310$, $p-value = 0.5518$

-- c.
State your decision and conclusion for the given problem.

Because $p-value > 0.05$, we fail to reject $H_0$ and can say there is not significant evidence that the true mean salmonella levels of all the melons from the factory is greater than $0.3$ MPN/g.
