---
title: "STAT 308 -- In Class Exercise"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

For the problems in which calculations are needed, please include your R code with your answers, otherwise you would not be given full credit on the exam.

The *cars* dataset is automatically available in base R.  The dataset provides information on 50 randomly selected cars and information how long it takes the cars to come to a complete stop when travelling at a particular speed.  The variables in the dataset are *speed*, the speed (in mph) at which the car was travelling when it began to stop and *dist*, the distance (in feet) it took the car to come to a complete stop.

1.  Suppose someone comes to you and says if we know how fast the car was travelling when it started to stop, we can make better predictions about the stopping distance of the car.  We want to start by creating a linear model predicting stopping distance based on the cars' speed.

a. Create a scatterplot of speed vs. distance.  Determine if a linear model is valid for the given data.
```{r}
```

b. State the least squares regression line for speed vs. distance.
```{r}
```

c.  Interpret the intercept and the slope of the regression lines.  Comment on the validity of these interpretations.

d. What is the predicted stopping distance for a car that is travelling at 17 mph?  Calculate and interpret 90% confidence and prediction intervals for this prediction.
```{r}
```

e. Suppose now I wish to test whether or not the average stopping distance is linearly with the car's speed?  Perform this hypothesis test, stating the correct null and alternative hypotheses, the test statistic, the distribution of the test statistic under the null hypothesis, p-value, and decision and conclusion in the context of the problem.

```{r}
```

f. Determine if the assumption of homoscedasticity is violated.
```{r}
```

g. Determine if the assumption of normally distributed residuals is violated.
```{r}
```

