
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

- 1. Use the following incomplete ANOVA table to answer the following questions.

|     |df    |Sums of Squares  |Mean Square  |F Value  |Pr(>f)   |
|-----|:-----|:----------------|:------------|:--------|:--------|
|Model|1 |             | | | 0.0011 |
|Error||  |10.5  |         |         |
|Total|60| |             |         |         |   

-- a. What are the error degrees of freedom?
```{r}
```

-- b. What are the sum of squared errors?
```{r}
```

-- c. What is the F statistic?
```{r}
```

-- d. What is the mean squares for the model?
```{r}
```

-- e. What are the sum of squares for the model?
```{r}
```

-- f. What is the r-squared for this dataset?
```{r}
```

-- g. What is the total variance for the response variable $Y$ ($s_Y^2$)?
```{r}
```

- 2. Reconsider the dataset cars in R.

-- a. Obtain an ANOVA table for this model.  
```{r}
```

-- b. Using the ANOVA table, perform an F test for a linear relationship between the car's speed and stopping distance.  Be sure to properly state your hypotheses, your test statistic, p-value, and a decision and conclusion at $\alpha = 0.05$.
```{r}
```

-- c. What is the distribution the test statistic follows under $H_0$?

-- d. Using the ANOVA table, calculate the value of $r^2$.  Interpret this value in the context of the problem.
```{r}
```

-- e. Determine the value of the correlation coefficient for the linear model.
```{r}
```


