
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

- 1. Use the following incomplete ANOVA table to answer the following questions.

|     |df    |Sums of Squares  |Mean Square  |F Value  |Pr(>f)   |
|-----|:-----|:----------------|:------------|:--------|:--------|
|Model|1 |             | | | 0.0011 |
|Error||  |10.5  |         |         |
|Total|60| |             |         |         |   

-- a. What are the error degrees of freedom?
```{r}
df_total <- 60
df_model <- 1
df_error <- df_total - df_model
df_error
```
$df_e$ = 59

-- b. What are the sum of squared errors?
```{r}
## MSE = SSE/df_error
MSE <- 10.5
SSE <- MSE*df_error
SSE
```
$SSE = 619.5$

-- c. What is the F statistic?
```{r}
# Use the p-value to get the F statistic
p_val <- 0.0011
f <- qf(0.0011,df1=df_model,df2=df_error,lower.tail=FALSE)
f
```
$F = 11.78$

-- d. What is the mean squares for the model?
```{r}
# F = MSM/MSE
MSM <- f*MSE
MSM
```
$MSM = 123.7$

-- e. What are the sum of squares for the model?
```{r}
# MSM = SSM/df_model
SSM <- MSM * df_model
SSM
```
$SSM = 123.7$

-- f. What is the r-squared for this dataset?
```{r}
SST <- SSM + SSE
r2 <- SSM/SST
r2
```
$r^2 = 0.1664$

-- g. What is the total variance for the response variable $Y$ ($s_Y^2$)?
```{r}
# sy2 = SST/df_total
sy2 <- SST/df_total
sy2
```
$s_y^2 = 12.39$

- 2. Reconsider the dataset cars in R.

-- a. Obtain an ANOVA table for this model.  
```{r}
mod <- lm(dist ~ speed,cars)
anova(mod)
```

-- b. Using the ANOVA table, perform an F test for a linear relationship between the car's speed and stopping distance.  Be sure to properly state your hypotheses, your test statistic, p-value, and a decision and conclusion at $\alpha = 0.05$.

Reduced Model: $\hat{Dist} = \beta_0$

Full Model: $\hat{Dist} = \beta_0 + \beta_1 Speed$

$$H_0: \beta_1 = 0$$

$$H_a: \beta_1 \neq 0$$
$F = 89.57$, $p = 1.49 \times 10^{-12}$, $p < \alpha$, so we reject $H_0$ and we can conclude there is a statistically significant linear relationship between the car's speed and its stopping distance.

```{r}
anova(mod)
```

-- c. What is the distribution the test statistic follows under $H_0$?

$F$ follows and F distribution with 1 and 48 degrees of freedom.

-- d. Using the ANOVA table, calculate the value of $r^2$.  Interpret this value in the context of the problem.
```{r}
SSM <- 21186
SSE <- 11354
SST <- SSM + SSE
r2 <- SSM/SST
r2
```
$r^2 = 0.6511$.  65.11% of the variation in car's stopping distance is explained through its linear relationship with the car's speed.

-- e. Determine the value of the correlation coefficient for the linear model.
```{r}
plot(dist ~ speed,cars)
# Look at the estimate of the slope of the least squares regression line
summary(mod)
# The estimate for this linear model is positive, so r is also positive
r <- sqrt(r2)
r
```

$r = 0.8069$.
