---
title: "Logistic Regression"
author: "Matt Stuart"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Regression

**Generalized Linear Model**: An extension of the linear regression methods discussed throughout the semester, without the need for the homoskedasticity or normality assumptions

- Linear Regression, Polynomial Regression, Logistic Regression, etc. are all examples of generalized linear models.

We will discuss ways to interpret logistic regression models using the Titanic dataset from the course webpage.

*Note*: Survived = 1 means the passenger survived the crash, and Survived = 0 means they did not

*Note*: There was room on the boat for Jack!

```{r}
Titanic <- read.csv("../data/Titanic.csv")
Titanic <- na.omit(Titanic) #Remove any rows with missing data
```

- We believe that age, sex, and passenger class are the most important explanatory variables that can be used to predicted whether or not a passenger survived.

- Let's fit a logistic regression model for the Titanic dataset where we want to predict survival rate of passengers by their age, sex, and passenger class.

```{r}
# glm is the function we use for a generalized linear model
# glm by default assumes traditional linear regression
# to use logistic regression, we have to specific a different family of distribution
# specifically, we use the binomial distribution
# R by default uses the logit as the activation/link function
mod <- glm(Survived ~ Sex + Age + as.character(Pclass), # turning Pclass to a categorical variable
           data = Titanic,
           family = "binomial")
summary(mod)
```
- Estimated logistic model:

$logit(\hat{p}) = log(\frac{\hat{p}}{1 - \hat{p}}) = 3.777 - 2.523 Male - 0.03670Age - 1.310 Class2 - 2.581 Class3$

## Interpretation Logisitic Regression Estimates

- The model parameters should always be interpreted in relationship to the **log odds** of the predicted probability.

### Questions

- Interpret the parameter estimate associated with Male in the context of the problem.

When a passenger is a male instead of a female, the predicted log odds of their survival of the Titanic decreases by 2.523, holding age and passenger class constant. (No units on the change in the log odds).

- Interpret the parameter estimate associated with Age in the context of the problem.

For a one year increase in age, the predicted log odds of their survival of the Titanic decreases by 0.0367, holding gender and passenger class constant.

- Predict the survival rate for a female passenger in class 1 aged 35.
```{r}
# Can use predict function for glms as well
newdata <- data.frame(Age = 35,Pclass = 1, Sex = "female")
predict(mod,newdata)
# The predict function gives us the log odds prediction
# Need to input into a different function to get probability
# I will show you how to write your own function in R
invlogit <- function(x){exp(x)/(1 + exp(x))}
logodds <- predict(mod,newdata)
invlogit(logodds)
```
The predicted probability that a 35 year old woman in Class 1 survived the Titanic crash is 0.9229.

## Hypothesis Testing in Logistic Regression

- We cannot use the t-test or the F-test because we do not assume the data are normally distributed

- We will use likelihoods to perform hypothesis tests

**Log-likelihoods**: An evaluation of how likely the value of a regression parameter is for the given set of data

- Related to the SSE in linear regression analysis

- The larger the log-likelihood, the better our generalized linear model is at making predictions

In the summary of a glm, the output gives us two deviance values:

- **Null Deviance**: -2*loglikelihood for the glm with no explanatory variables ($df_{Null} = n - 1$)

- **Residual Deviance**: - 2*loglikelihood for the glm with the given explanatory variables ($df_{res} = n - p$, $p$ is the number of parameters in the model)

