---
title: "STAT 308 -- Chapter 6"
output: 
  pdf_document
header-includes:
  - \usepackage{framed}
---

```{r setup, include = FALSE}
# This line of code tells the document all the display defaults

knitr::opts_chunk$set(echo = TRUE, include = TRUE, fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

# Background Information

We have previously discussed how we can model observed values of $Y$ by our knowledge of the \textbf{independent variable} $X$.  We discussed methods to assess if $\beta_1 > 0$ and create confidence intervals for $\beta_1$ and $\mu_{Y|X}$ and prediction intervals for $Y$.  We also discussed \textbf{graphical methods} to asssess the goodness of fit of our linear model.  We will now discuss \textbf{numerical methods} to make inference on our linear model.

# Important Definitions

\begin{framed}
\textbf{Correlation Coefficient:} A number describing the \textit{strength} and \textit{direction} of the \textit{linear association} between $X$ and $Y$.   
\end{framed}

The sample correlation coefficient is defined as $$r = \frac{\sum_{i=1}^n(x_i-\bar{x})(y_i - \bar{y})}{\left(\sum_{i=1}^n(x_i - \bar{x})^2\sum_{i=1}^n(y_i - \bar{y})^2\right)^{1/2}} = \frac{s_{xy}}{s_xs_y},$$ where $s_{xy}$ is the sample \textbf{covariance} between the observed $x$ and $y$.

It can be noted that $r$ is directly related to the estimated regression slope $\hat{\beta}_1$, $$r = \frac{s_x}{s_y}\hat{\beta}_1.$$

## Example:

Recall the \texttt{bloodpressure} dataset we have previously used.  Calculate the correlation coefficient between Age and Systolic Blood Pressure.

```{r}
library(tidyverse)
```

## Properties of the sample correlation coefficient

- $r$ is a value between -1 and 1.

-- $r = 1$ means there is a direct positive linear relationship between $X$ and $Y$.

-- $r = -1$ means there is a direct negative linear relationship between $X$ and $Y$.

- $r$ is a \textit{unitless} measure

- $r$ has the same sign as $\hat{\beta}_1$.  That is $$r > 0 \longleftrightarrow \hat{\beta}_1 > 0$$ and $$r < 0 \longleftrightarrow \hat{\beta}_1 < 0$$

- $r = 0$ means there is no \textbf{linear} association between $X$ and $Y$.  That does not mean there is no pattern at all that can be made out by the graph of $X$ and $Y$.

### Examples of scatterplots and the correlation coefficient

```{r, echo=FALSE}
set.seed(0604001)
x <- rnorm(n=50,mean=0,sd=1)
y_pos <- rnorm(n=50,mean=x,sd=1)
y_neg <- rnorm(n=50,mean=-x,sd=1)
y_par <- rnorm(n=50,mean=10*x^2-x^3,sd=1)
y_non <- rnorm(n=50,mean=0,sd=1)
```

```{r, echo=FALSE}
data.frame(x=x,y=y_pos) %>% 
  ggplot(aes(x=x,y=y)) + 
  geom_point()
paste0("Correlation between X and Y is ",round(cor(x,y_pos),2),".")
```

```{r, echo=FALSE}
data.frame(x=x,y=y_neg) %>% 
  ggplot(aes(x=x,y=y)) + 
  geom_point()
paste0("Correlation between X and Y is ",round(cor(x,y_neg),2),".")
```

```{r, echo=FALSE}
data.frame(x=x,y=y_non) %>% 
  ggplot(aes(x=x,y=y)) + 
  geom_point()
paste0("Correlation between X and Y is ",round(cor(x,y_non),2),".")
```

```{r, echo=FALSE}
data.frame(x=x,y=y_par) %>% 
  ggplot(aes(x=x,y=y)) + 
  geom_point()
paste0("Correlation between X and Y is ",round(cor(x,y_par),2),".")
```

### Example

What is the correlation coefficient of the blood pressure dataset?

\begin{framed}
\textbf{Bivariate Normal Distribution:} A distribution that describes the joint relationship between two different normally distributed random variables $X$ and $Y$
\end{framed}

## Parameters of Bivariate Normal Distribution:

- $\mu_X$: univariate mean of $X$

- $\mu_Y$: univariate mean of $Y$

- $\sigma_X^2$: univariate variance of $X$

- $\sigma_Y^2$: univariate variance of $Y$

- $\rho_{XY}$: correlation between $X$ and $Y$

A nice property of the bivariate normal distribution is that we can slice the distribution at a fixed value of $X$ to obtain the \textit{conditional distribution} of $Y$ at a given value of $X$.  This distribution is also normally distributed with 

- $\mu_{Y|X} = \mu_Y + \frac{\rho_{XY}\sigma_Y}{\sigma_X}(X - \mu_X)$ and 

- $\sigma_{Y|X}^2 = \sigma_Y^2(1 - \rho_{XY}^2).$

Recall from the Chapter 5 and 6 notes, we can say that

- $\hat{\beta}_1 = \frac{rs_y}{s_x}$ and 

- $\hat{\beta}_0 = \bar{y} - \hat{\beta}_1\bar{x}.$

If we substitute in the parameters for their respective estimates, we have

- ${\beta}_1 = \frac{\rho_{XY}\sigma_Y}{\sigma_X}$ and 

- ${\beta}_0 = \mu_{y} - {\beta}_1\mu_{x}.$

Then, using some substitution, we have

- $\mu_{Y|X} = \beta_0 + \beta_1 X,$

showing the relationship between the least squares regression line and the bivariate normal distribution!

Now, if we were to take the formula for $\sigma_{Y|X}^2$ and solve for $\rho_{XY}^2$, we would get $$\rho_{XY}^2 = \frac{\sigma_{Y}^2 - \sigma_{Y|X}^2}{\sigma_{Y}^2}.$$

In other words, $\rho_{XY}^2$ is the

## R-squared

\begin{framed}
\textbf{R-Squared}($r^2$): the percent of variation in the response variable $Y$ that can be explained through its linear relationship with the explanatory variable $X$
\end{framed}

Formally, $$r^2 = \frac{SSY-SSE}{SSY},$$ where $SSY = \sum_{i=1}^n (y_i - \bar{y})^2$ and $SSE = \sum_{i=1}^n(y_i - (\hat{\beta}_0 + \hat{\beta}_1x_i))^2$.

### Example

Find the $r^2$ of the systolic blood pressure dataset.  Interpret this value in the context of the given problem.

```{r}
```

### About R-squared

- Naturally, $r^2$ is the square of the sample correlation coefficient, so $0 \leq r^2 \leq 1$

- The larger the value of $r^2$, the more variance in $Y$ we can explain through its linear relationship with $X$, and thus, the stronger the linear relationship between the two variables

-- If $r^2 = 1$, all of the variation in $Y$ can be explained linearly by $X$ (in other words, $SSE = 0$)

-- If $r^2 = 0$, no variation in $Y$ can be explained linearly by $X$

- $r^2$ does NOT measure the magnitude of $\hat{\beta}_1$ (i.e. $r^2$ can be close to one, but $\hat{\beta}_1$ may still be close to zero, or $r^2$ can be close to zero, but $\hat{\beta}_1$ may be large.)

```{r}
```

- $r^2$ is NOT a measure of the appropriateness of the linear model.





