---
title: "STAT 308 -- Take Home Exam"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

This take home exam is worth a total of 100 points.  For the problems in which calculations are needed, please include your *R* code with your answers.  Turn in this exam to Sakai by Friday, February 24 at **3:35 pm**.

1.  The following questions will test your basic *R* skills.

a. [4 pts] Construct a vector of the numbers 0,4,3,6,1,6,2,4,1,8,9,5.  Calculate the mean, median, and variance of the vector of numbers.

b. [2 pts] Suppose $X \sim \mathcal{N}(2,3)$.  Calculate $P(X > 4)$.

c. [2 pts] Suppose $X \sim t_{df=9}$.  Calculate the $80^{th}$ percentile of $X$.

d. [2 pts] Suppose $X \sim F_{df1=2,df2=12}$.  Calculate $P(X < 2)$.

2.  For a random sample of $n = 50$, we obtain the following sample statistics:
$$\bar{x} = 4, \bar{y} = 7, s_x = 1.3, s_y = 2.9, r = 0.8.$$

a. [4 pts] What is the estimate of the least squares regression line?
```{r}
```

b. [2 pts] What is the percent variation in $Y$ that can be explained by its linear relationship with $X$?
```{r}
```

c. [4 pts] What are the sum of squared errors for the linear model of $Y$ with $X$ as an explanatory variable?
```{r}
```

d. [3 pts] What are the mean squares for the model?
```{r}
```

e. [3 pts] What is the estimate of the regression variance?
```{r}
```

f. [2 pts] Calculate an appropriate test statistic in a hypothesis test for a significant linear relationship between $X$ and $Y$.
```{r}
```

g. [2 pts] What is the distribution of the test statistic in (f) under $H_0$?

3. Consider the following incomplete ANOVA table for a simple linear regression model.

|     |df    |Sums of Squares  |Mean Square  |F Value  |Pr(>f)   |
|-----|:-----|:----------------|:------------|:--------|:--------|
|Model|1 |              | | 32.7 |  |
|Error||  | 48.2 |         |         |
|Total|59|  |             |         |         |  

a. [2 pt] What are the error degrees of freedom?
```{r}
```

b. [3 pts] What are the sums of squares for the model and the sum of squared errors?
```{r}
```

c. [3 pts] What is the p-value used to test for a significant linear relationship between $X$ and $Y$?
```{r}
```

d. [2 pts] What is the sample variance of $Y$ ($s_y^2$)?
```{r}
```

e. [2 pts] What is the r-squared for the simple linear regression model?
```{r}
```

f. [3 pts] If I were to construct a 95% confidence interval for the slope, $\beta_1$, would 0 be inside the interval?  Briefly explain your answer.

4.  Suppose I create a simple linear regression model for a particular dataset.  From there, I calculate 95% confidence and prediction intervals for the value of $Y$ when $X = 4.2$ and when $X = 3.5$ (the sample mean of $X$).  The four intervals are (70.19,71.60), (80.81,82.70), (59.23,82.56), and (70.07,93.44).

a. [2 pts] Which of the four intervals is the 95% confidence interval for $\hat{Y}$ when $X = 4.2$?

b. [2 pts] Which of the four intervals is the 95% prediction interval for $\hat{Y}$ when $X = 4.2$?

c. [1 pt] Which of the four intervals is the 95% prediction interval for $\hat{Y}$ when $X = 3.5$?

The dataset *realestate.csv* contains information on a sample of 414 houses that were recently sold in New Taipei City, Taiwan.  The variables listed are

- *Age*: The age of the house in years

- *MRT*: The distance (in meters) to the nearest MRT (public transportation) station

- *Convenience*: The number of convenience stores within 1 kilometer of the house

- *Latitude*: The degrees latitude of the location of the house

- *Longitude*: The degrees longitude of the location of the house

- *Price*: The house's sales price (in millions of New Taiwan dollars).

5. Suppose we wish to create a model to predict the house's sales price using the distance to the nearest MRT station as an explanatory variable.

a. [2 pts] Create a scatterplot of the house's sales price against distance to the nearest MRT station.  Comment on the validity of a linear relationship.
```{r}
```

b. [3 pts] Calculate a least squares regression line using *R*.  Formally state the regression line.
```{r}
```

c. [3 pts] Interpret the estimate of the slope in the context of the problem.

d. [3 pts] Interpret the estimate of the intercept in the context of the problem.  Does this interpretation make sense?  Why or why not?

e. [2 pts] Report and interpret the r-squared in the context of the problem.
```{r}
```

f. [4 pts] Calculate a 95% confidence interval for the slope.  Interpret this interval in the context of the problem.
```{r}
```

g. [5 pts] Perform a hypothesis test for a linear relationship between distance to the nearest MRT station and home sales price.  Be sure to formally state all the necessary information to perform a formal hypothesis test.  Assume $\alpha = 0.05$.
```{r}
```

h. [4 pts] Calculate a prediction as well as a 95% prediction interval for the sales price of a house that is 500 meters from the nearest MRT station.  Interpret the prediction interval in the context of the given problem.
```{r}
```

i. [2 pts] Determine if the assumption of heteroscedasticity is violated.
```{r}
```

j. [2 pts] Determine if the assumption of normally distributed residuals is violated.
```{r}
```

6. Suppose we now wish to create a multiple linear regression model for sales price with all of the other variables in the dataset as explanatory variables.

a. [5 pts] Report the least squares regression line for this linear model.
```{r}
```

b. [3 pts] Interpret the parameter associated with the explanatory variable *MRT* in the context of the problem.

c. [3 pts] Interpret the parameter associated with the explanatory variable *Age* in the context of the problem.

d. [2 pts] Report the r-squared for the linear model.  Why is this value of r-squared larger than the r-squared you reported in 5(e)?
```{r}
```

e. [3 pts] What are the error degrees of freedom and the mean squared error for this linear model?
```{r}
```

f. [4 pts] Calculate and interpret a 90% confidence interval for a home's predicted sales price for a house that is 10 years old, 1000 meters from the nearest MRT station, has 4 convenience stores within 1 km, and is located at 24.98 degrees latitude and 121.53 degrees longitude.
```{r}
```
