

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE,
fig.align="center",warnings = FALSE,fig.width = 5,fig.height=3)
```

This take home exam is worth a total of 100 points.  For the problems in which calculations are needed, please include your \texttt{R} code with your answers.  Turn in this exam to Sakai by Thursday, November 17 at 2:30 pm.

1.  [3 pts each] For each of the following problems with a sample size of $n$, calculate an appropriate test statistic and p-value for a linear model with $p-1$ total explanatory variables in the model. In other words, we have $p$ parameters in the linear model (one for each of the $p-1$ explanatory variables, and the intercept).

a. $n = 50$, $p = 4$, $SSM = 150.32$, $SSE = 230.44$.
```{r}
```

b. $n = 15$, $p = 3$, $SSM = 77.3$, $SSE = 125.44$.
```{r}
```

c. $n = 60$, $p = 6$, $SSM = 50.32$, $SSE = 250.78$.
```{r}
```

d. $n = 75$, $p = 5$, $SSM = 3.22$, $SSE = 9.87$.
```{r}
```

e. $n = 28$, $p = 2$, $SSM = 60.44$, $SSE = 514.32$.
```{r}
```

2. Use the below degrees of freedom and sums of squares from ANOVA tables from a multiple linear regression with three explanatory variables to answer the following questions.

- Model 1: $X_1$ as only explanatory variable

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |4 | 2341.53 |
|Error |95 | 1032.35 |

- Model 2: $X_1$ and $X_2$ as explanatory variables

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |5 | 2609.87 |
|Error |94 |764.01 |

- Model 3: $X_1$, $X_2$, and $X_3$ as explanatory variables

|     |df    |Sums of Squares  |
|-----|:-----|:----------------|
|Model |6 | 2701.22 |
|Error |93 |672.66 |

a. [2 pts] Is $X_1$ a categorical variable?  If it is, how many possible values are there for $X_1$?

b. [2 pts] What is the sample size $n$ for the dataset?

c. [2 pts] What are the total sums of squares for the response variable $Y$?

d. [2 pts] What is the mean squared error for the model with $X_1$ and $X_2$ included.
 
e. [2 pts] What is the $r^2$ for the model with $X_1$, $X_2$, and $X_3$ included?

f. [2 pts] What is the additional explained variation in the response when we add $X_3$ to a model that already includes $X_1$ and $X_2$?

g. [2 pts] What are the additional model sums of squares when we add $X_2$ and $X_3$ to a model that includes $X_1$?

h. [2 pts] What are the model and error degrees of freedom for testing if including $X_1$, $X_2$ and $X_3$ in a model is statistically significant.

i. [2 pts] Suppose we want to test if adding $X_3$ to a model that already includes $X_1$, $X_2$ is statistically significant.  What are the model and error degrees of freedom?

j. [4 pts] Find an F statistic and p-value for testing if adding $X_2$ to a model that already includes $X_1$ is statistically significant.

3.  For this problem, consider the mtcars dataset available in base R.  Use the command "?mtcars" to obtain information about the overall dataset, as well as the variables contained within it.  

a. Suppose we want to fit a linear model to predict a car's fuel efficiency by 

- Number of Cylinders

- Horsepower

- Weight

- Transmission

- Quarter Mile Time

i. [4 pts] Is there a potential issue with multicollinearity with this dataset?  If so, explain why and what variable should we consider removing.
```{r}
```

ii. [4 pts] Perform a formal hypothesis test to determine if the explanatory variable with the highest VIF in the linear model in (a) is statistically significant if all of the other explanatory variable are included in the model.  Assume $\alpha = 0.05$.
```{r}
```

iii. [2 pts] Is there still an issue of multicollinearity after the variable with the highest VIF is removed?  Explain your answer.
```{r}
```

b. [2 pts] Check if the assumptions of homoskedasticity and normally distributed are violated for the original model in (a).  If they are not met, make a suggestion for how we can adjust our model.

```{r}
```

c. Suppose that, after discussion with fellow researchers, we decide to transform the response to log-fuel efficiency, and fit a linear model for log-fuel efficiency with all of the other variables still included.

- i. [3 pts] Write the least squares regression line.
```{r}
```

- ii. [3 pts] Interpret the parameter associated with number of cylinders in the context of the problem.

- iii. [3 pts] Interpret the parameter associated with transmission in the context of the problem.

- iv. [4 pts] What do we predict the fuel efficiency to be for a car with 6 cylinders, 130 horsepower, weighs 2200 lbs, is a manual transmission, and has an 18 second quarter-mile time.

- v. [3 pts] Interpret the $r^2$ for this model in the context of the problem.
```{r}
```

- vi. [4 pts] Perform a formal hypothesis test to see if our transformed linear model is statistically significant.  Assume $\alpha = 0.05$.
```{r}
```

- vii. [2 pts] Determine if the assumptions for homoskedasticity and normality are violated for the transformed model.
```{r}
```

4.  Consider the \texttt{lifeexp} dataset on the course webpage which contains information on countries from the year 2015.  The variables contained in the dataset are defined as follows:

- \texttt{Status}: Development status of the country

- \texttt{Life.Expectancy}: Country's average life expectancy in years

- \texttt{Adult.Mortality}: Number of deaths of people between 15 and 60 per 1000 population

- \texttt{infant.deaths}: Number of infant deaths per 1000 population

- \texttt{Measles}: Number of measles cases per 1000 population

- \texttt{BMI}: average BMI for entire population

- \texttt{under.five.deaths}: Number of Under five deaths per 1000 population

- \texttt{Polio}: Polio immunization coverage for 1 year olds

- \texttt{Diptheria}: Diphtheria tetanus toxoid and pertussis immunization coverage for 1 year olds

- \texttt{HIV.AIDS}: Deaths per 1000 live births from HIV/AIDS (0-4 year olds)

- \texttt{Income.composition.of.resources}: Human Development Index in terms of income composition of resources (from 0 to 1)

- \texttt{Schooling}: Number of Years of Schooling

We are interested in creating a regression model that predicts a country's life expectancy based on the other factors in the dataset.

a.  Suppose we know that both adult mortality and income composition of resources are significant in a linear model that predicts life expectancy.  We would like to know if adding development status signficantly improves the predictive ability of our model.

- i. [3 pts] Report the least squares regression lines for predicting life expectancy for both developed and developing countries with adult mortality and income composition of resources included.
```{r}
```

- ii. [3 pts] What is the expected difference in life expectancy between developed and developing countries for fixed levels of adult mortality and income composition of resources?

- iii. [4 pts] Perform a formal hypothesis test to determine if adding development status to our linear model significantly improves our model's predictive ability for life expectancy.  Be sure to include all pieces of information needed to perform a hypothesis test.  Assume $\alpha = 0.05$.

```{r}
```

- iv. [5 pts] Perform a formal hypothesis test to determine if EITHER the linear relationship between adult mortality and life expectancy OR the linear relationship between income composition of resources and life expectancy is different for the two different development statuses.  Be sure to include all pieces of information needed to perform a hypothesis test.  Assume $\alpha = 0.05$.

```{r}
```

b.  We now would like to determine the "best" overall prediction model for life expectancy using all of the other variables in our dataset as predictors.

- i. [3 pts] Should we include any polynomial terms in our model? Explain your reasoning.
```{r}
```

- ii. [6 pts]  Assume we include no polynomial terms or interaction terms.  Use backward selection with AIC as your decision criteria to eliminate variables from your model.  Be sure to state which variables were eliminated at each step.
```{r}
```

- iii. [3 pts] Write your final regression model for predicting life expectancy.
```{r}
```

- iv. [2 pts] Report the $r^2$ and mean squared error for your final linear model.
```{r}
```

